---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# estimators <img src=`r file.path("man", "figures", "logo.png")` align="right" height="139" alt="logo"/>

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/estimators)](https://CRAN.R-project.org/package=estimators)
[![R-CMD-check](https://github.com/thechibo/estimators/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/thechibo/estimators/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/thechibo/estimators/branch/main/graph/badge.svg)](https://app.codecov.io/gh/thechibo/estimators?branch=main)
<!-- badges: end -->

## Introduction

The `estimators` R package performs parameter estimation in common distribution families, making moment, maximum likelihood, and state-of-the-art estimators more accessible.

## Key Features

- Point Estimation: The package offers a comprehensive selection of moment-type estimators, allowing users to quickly and easily estimate distribution parameters from their data.
- Asymptotic Variance Matrices: In addition to parameter estimates, the package computes and provides asymptotic variance-covariance matrices.
- Computational Efficiency: The `estimators` R package is designed with computational efficiency in mind. All algorithms are optimized to handle large datasets with minimal computation time, ensuring that users can perform parameter estimation even with substantial data volumes.

## Installation

You can install the development version of `estimators` from github by running the following line of code:

```{r, eval=FALSE}
 devtools::install_github("thechibo/estimators")
```

More details can be found in the [estimators Github repository](https://github.com/thechibo/estimators "estimators Github repository").

## Documentation

Detailed documentation, along with reproducible examples, can be found in the package vignette `vignette(topic = "estimators", package = "estimators")`.

## Team

The `estimators` package is developed in the 
[Mathematics Department](https://en.math.uoa.gr/ "Mathematics Department Homepage") of the [University of Athens](https://en.uoa.gr/ "University of Athens Homepage"). The package maintainer is [Ioannis Oikonomidis](http://users.uoa.gr/~goikon/ "Ioannis Oikonomidis Homepage"), working under the supervision of [Prof. Samis Trevezas](http://scholar.uoa.gr/strevezas/ "Samis Trevezas Homepage").
